NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2019-02-05 Manuscript/manuscriptPS.tex
Style edits to MD methods.

\subsection{C--H bond order parameters from the natural abundance $^{13}$C\,NMR}

Headgroup and glycerol backbone C--H bond order parameters of POPS
were determined from the chemical-shift resolved dipolar splittings
measured with a R-type Proton Detected Local Field (R-PDLF) experiment~\cite{dvinskikh04}.
The corresponding order parameter signs were measured with a S-DROSS experiment~\cite{gross97}
using natural abundance $^{13}$C solid state NMR spectroscopy as described previously \cite{ferreira13,ferreira16}.
The experiments were done in a Bruker Avance III 400 spectrometer operating at a $^1$H Larmor frequency of 400.03 MHz.
Magic angle spinning (MAS) of the sample was used at a frequency of 5.15 kHz (R-PDLF experiment) and 5 kHz (S-DROSS experiment).
The following experimental setups were used.

{\emph{C--H bond order parameters from the R-PDLF experiment.}} The parameters are described according to Figures 1c and 2c of the original reference
for the R-PDLF experiment~\cite{dvinskikh04}.  The refocused-INEPT delays were $\tau_1=$\,1.94\,ms and $\tau_2=$\,0.97\,ms.
Radio frequency pulses with the nutation frequencies: 46.35 kHz (R18$^7_1$ pulses), 63.45 kHz ($^{13}$C 90$^{\rm{o}}$ and 180$^{\rm{o}}$),
50 kHz (SPINAL64 $^1$H decoupling pulses).
\todo{Verb missing. Would it be correct to say: 'Radio frequency pulses had the nutation frequencies:'?}
The $t_1$ increment was equal to 10.79 $\mu$s $\times18\times2$, and 32 points in the indirect
dimension were recorded using 1024 scans for each, with a recycle delay of 5\,s and a spectral with \todo{'with' $\to$ 'width'?} of 149.5 ppm.

\emph{Order parameter signs from the S-DROSS experiment.}
The parameters are described according to Figures 1b and 1c of the original reference for the S-DROSS
experiment~\cite{gross97}. The refocused-INEPT delay $\delta_2$ was 1.19 ms. The $\tau_1$ and $\tau_2$ in the S-DROSS recoupling
blocks $R$ were set as $\tau_1=$\,39.4\,$\mu$s and $\tau_2=$\,89.4 $\mu$s. Radio frequency pulses with the nutation
frequencies: 63.45 kHz ($^{13}$C 90$^{\rm{o}}$ and 180$^{\rm{o}}$), 50 kHz ($^1$H\,SPINAL64 decoupling).
\todo{Verb missing. Would it be correct to say: 'Radio frequency pulses had the nutation frequencies:'?}
The $t_1$ increment (dipolar recoupling dimension) was 800 $\mu$s, and a total of 8 points along $t_1$ were
measured using 1024 scans for each, with a recycle delay of 5\,s and a spectral with \todo{'with' $\to$ 'width'?}  of 149.5 ppm.

\emph{Numerical simulations of S-DROSS curves.}
The numerical simulations of S-DROSS curves were performed with the SIMPSON simulation package \cite{bak00}
using as input the $^{13}$C--$^1$H
dipolar couplings, either as determined by the R-PDLF experiments, or as calculated from the known $^2$H quadrupolar couplings \cite{browning80}.
The chemical shift anisotropy and homonuclear couplings were neglected, and the input file {\it{rep2000}} was used to simulate a random
distribution of bilayer orientations in the samples studied.

\emph{Sample preparation.}
The sample was prepared simply by mixing the POPS \todo{From some company?} with water (lipid:water 60:40\,wt-\%) in an Eppendorf
tube by mixing and centrifuging the sample repeatedly until a \todo{visually(?)} homogeneous viscous fluid was obtained.
Then 20 mg of the sample was  transferred to an NMR insert suitable for 4 mm NMR rotors.  
\todo{Maybe we need little bit more information about the mixing procedure?}

\subsection{Molecular dynamics simulations}
Molecular dynamics simulation data was collected using
the Open Collaboration method \cite{botan15}.
The NMRlipids project blog (\url{www.nmrlipids.blogspot.fi}) and
the GitHub repository (\url{www.github.com/NMRLipids/NMRlipidsIVotherHGs})
were used as the communication platforms.
The simulated systems are listed in 
Tables \ref{PSsystems} (pure PS systems without additional ions) 
and \ref{mixedIONsystems} (mixed PC:PS systems with various ions concentrations).
Further simulation details are given in the SI.
The simulation data is also indexed in the
searchable database (\url{www.nmrlipids.fi}),
and in the NMRlipids/MATCH GitHub repository (\url{https://github.com/NMRLipids/MATCH}).

The C-H bond order parameters were calculated directly
from the definition
\begin{equation}
S_{\rm CH}=\frac{1}{2}\langle 3\cos^2\theta -1 \rangle,
\end{equation}
where $\theta$ is the angle between the C-H bond and the membrane normal.
Angular brackets point to the average over all sampled configurations.
The order parameters were first calculated averaging over time separately
for each lipid molecule in the system. The average and
the standard error of the mean were then calculated over different lipids.
The python program ({\it scripts/calcOrderParameters.py}) utilizing
MDAnalysis library \cite{agrawal11,gowers16} is available in Ref. \citenum{MATCHgit}. 
The number density profiles were calculated using {\it gmx density} tool
from Gromacs sofware package \cite{gromacsMANUAL}.

>>>>

\subsection{C-H bond order parameters from the natural abundance $^{13}$C NMR}

Headgroup and glycerol backbone C-H bond order parameters of POPS
were determined from the chemical-shift resolved dipolar splittings
measured with a R-type Proton Detected Local Field (R-PDFL) experiment~\cite{dvinskikh04}.
The corresponding order parameter signs were measured with a S-DROSS experiment~\cite{gross97}
using natural abundance $^{13}$C solid state NMR spectroscopy as described previously \cite{ferreira13,ferreira16}.
The experiments were done in a Bruker Avance III 400 spectrometer operating at a $^1$H Larmor frequency of 400.03 MHz.
Magic angle spinning (MAS) of the sample was used at a frequency of 5.15 kHz (R-PDLF experiment) and 5 kHz (S-DROSS experiment).
The following experimental setups were used.

{\emph{C-H bond order parameters from the R-PDLF experiment}}. The parameters are described according to Figures 1c and 2c of the original reference
for the R-PDLF experiment~\cite{dvinskikh04}.  The refocused-INEPT delays $\tau_1$ and $\tau_2$ were 1.94 ms and 0.97 ms, respectively.
Radio frequency pulses with the nutation frequencies: 46.35 kHz (R18$^7_1$ pulses), 63.45 kHz ($^{13}$C 90$^{\rm{o}}$ and 180$^{\rm{o}}$),
50 kHz (SPINAL64 $^1$H decoupling pulses). The $t_1$ increment was equal to 10.79 $\mu$s $\times18\times2$ and 32 points in the indirect
dimension were recorded using 1024 scans for each, with recycle delay of 5 s and a spectral with of 149.5 ppm.

\emph{Order parameter signs from the S-DROSS experiment}.
The parameters are described according to Figures 1b and 1c of the original reference for the S-DROSS
experiment~\cite{gross97}. The refocused-INEPT delay $\delta_2$ was 1.19 ms. The $\tau_1$ and $\tau_2$ in the S-DROSS recoupling
blocks $R$ were set as 39.4 $\mu$s and 89.4 $\mu$s, respectively. Radio frequency pulses with the nutation
frequencies: 63.45 kHz ($^{13}$C 90$^{\rm{o}}$ and 180$^{\rm{o}}$), 50 kHz ($^1$H SPINAL64 decoupling).
The $t_1$ increment (dipolar recoupling dimension) was 800 $\mu$s and a total of 8 points along $t_1$ were
measured using 1024 scans for each with a recycle delay of 5 s and a spectral with of 149.5 ppm.

\emph{Numerical simulations of S-DROSS curves.}
The numerical simulations of S-DROSS curves were performed with the SIMPSON simulation package \cite{bak00} using the $^{13}$C--$^1$H
dipolar couplings determined by the R-PDLF experiments or calculated from the known $^2$H quadrupolar couplings \cite{browning80} as input.
The chemical shift anisotropy and homonuclear couplings were neglected, and the input file {\it{rep2000}} was used to simulate the random
distribution of bilayer orientations in the samples studied.

\emph{Sample preparation}
The sample was prepared simply by mixing the POPS with water (lipid:water 60:40 wt\%) in an eppendorf
tube mixing and centrifuging the sample repeatedly until an homogeneous viscous fluid was obtained.
20 mg of sample was then transferred to an NMR insert suitable for 4 mm NMR rotors.  
\todo{Maybe we need little bit more information about the mixing procedure?}

\subsection{Molecular dynamics simulations}
Molecular dynamics simulation data were collected using
the Open Collaboration method \cite{botan15}, with
the NMR\-lipids Project blog (\url{nmrlipids.blogspot.fi}) and
GitHub repository (\url{github.com/NMRlipids/NMRlipidsIVotherHGs})
as the communication platforms.
The simulated systems are listed in 
Tables \ref{PSsystems} (pure PS bilayers without additional ions) 
and \ref{mixedIONsystems} (mixed PC:PS bilayers at various salt concentrations).
Further simulation details are given in the SI.
The simulation data are also indexed in a
searchable database available at \url{www.nmrlipids.fi},
and in the NMRlipids/MATCH repository (\url{github.com/NMRlipids/MATCH}).

The C--H bond order parameters were calculated directly
from the carbon and hydrogen positions using the definition
\begin{equation}
S_{\rm CH}=\frac{1}{2}\langle 3\cos^2\theta -1 \rangle,
\end{equation}
where $\theta$ is the angle between the C--H bond and the membrane normal
(taken to align with $z$, with bilayer periodic in the $xy$-plane).
Angular brackets denote average over all sampled configurations.
The order parameters were first calculated averaging over time separately
for each lipid in the system. The average and
the standard error of the mean were then calculated over different lipids.
Python program ({\tt scripts/calcOrderParameters.py}) that uses the
MDAnalysis library \cite{agrawal11,gowers16} is available in Ref. \citenum{MATCHgit}. 
The number density profiles were calculated using the {\tt gmx density} tool
of the Gromacs sofware package \cite{gromacsMANUAL}.
